Hurricane Evacuation Strategies for the State of Mississippi

### 2015-11-16 manuscript.tex
May need touchups

\section{Stochastic Model of Landfall-Avodiant Evacuations}
\label{sec:stochastic}
  \subsection{Model Description}
    \par In this model, instead of giving citizens particular evacuation plan, we give them a simple rule: move directly away from the hurricane landfall if possible, else move North or North-West. Using the statewide broadcast system, the state of Mississippi can tell its citizens where the expected landfall will be, and which counties will be the most affected. In which case, individual citizens are instructed to leave the counties that will be flooded, and to do so as outlined above. This means less coordination from the state, and more options for each citizens. At the cost of not having designated areas, and therefore designated refuge areas for citizens, evacuation is faster because it involves less per-county organization.\\
    \par As opposed to the model described in \nameref{sec:maxflow}, we also allow citizens to leave the state of Mississippi. For example, Arkansas is less likely to be hit by the full force of hurricanes that move through Mississippi \cite{5news}, as such, it is a good place to take refuge while the storm passes. In order to achieve this, we add ``unofficial'' nodes in the graph representation of Mississippi. Figure~\ref{fig:highway_map} shows that Hancock and Jackson Counties have roads that lead into Louisiana and Alabama respectively, both of which are potential routes away from the Hurricane's landfall, meaning that such nodes are important when movement can also be lateral. Because we are only interested in the movement of citizens within Mississippi, we can assume that citizens that reach the out-of-state nodes keep moving further and further away from the hurricane, though we are not tracking them. We created four out-of-state nodes North, South, East and West of Mississippi, each connected in a way that satisfies the highways mapped in Figure~\ref{fig:highway_map}.
  \subsection{Parameter Values and Justification}
    \par The parameters to this model are the same as the parameters for the models outlined above, except that this model uses the complete highway map of Mississippi, instead of only considering vertical highways. This algorithm expects three inputs outside of the parameters: the first, less accurate estimate of the affected counties, the more accurate second estimate and the final, most accurate estimate. Notice that as described, these inputs actually abstract away the hurricane's category. Instead, the category is reflected by the size of the list of the affected counties given as the respective inputs for the estimates. In order to generate these input, we used the historical hurricane data. Given the landfall and the hurricane's category, we can generate three sets of decreasing size containing neighboring counties. We can then run the simulation on these inputs.
  \subsection{Results}
    This model's main goal is to find the amount of time it takes to evacuate the affected states.
    \begin{figure}
      \center
      \begin{subfigure}[b]{0.5\textwidth}
        \center
        \includegraphics[width=\linewidth]{figures/pred_hancock-crop.pdf}
        \caption{Predicted Evacuation for Landfall in Hancock County}
      \end{subfigure}~
      \begin{subfigure}[b]{0.5\textwidth}
        \center
        \includegraphics[width=\linewidth]{figures/pred_jackson-crop.pdf}
        \caption{Predicted Evacuation for Landfall in Jackson County}
      \end{subfigure}
      \caption{}
      \label{}
    \end{figure}
  \subsection{Strengths and Weaknesses}

>>>>

\section{Heuristic Model of Landfall-Avodiant Evacuations}
\label{sec:heuristic}
    \par In this model, instead of giving citizens particular evacuation plan, we give them a simple rule: move directly away from the hurricane landfall if possible, else move North or North-West. Using the statewide broadcast system, the state of Mississippi can tell its citizens where the expected landfall will be, and which counties will be the most affected. In which case, individual citizens are instructed to leave the counties that will be flooded, and to do so as outlined above. This means less coordination from the state, and more options for each citizens. At the cost of not having designated areas, and therefore designated refuge areas for citizens, evacuation is faster because it involves less per-county organization.\\
    \par As opposed to the model described in \nameref{sec:maxflow}, we also allow citizens to leave the state of Mississippi. For example, Arkansas is less likely to be hit by the full force of hurricanes that move through Mississippi \cite{5news}, as such, it is a good place to take refuge while the storm passes. In order to achieve this, we add ``unofficial'' nodes in the graph representation of Mississippi. Figure~\ref{fig:highway_map} shows that Hancock and Jackson Counties have roads that lead into Louisiana and Alabama respectively, both of which are potential routes away from the Hurricane's landfall, meaning that such nodes are important when movement can also be lateral. Because we are only interested in the movement of citizens within Mississippi, we can assume that citizens that reach the out-of-state nodes keep moving further and further away from the hurricane, though we are not tracking them. We created four out-of-state nodes North, South, East and West of Mississippi, each connected in a way that satisfies the highways mapped in Figure~\ref{fig:highway_map}.
    \par The parameters to this model are the same as the parameters for the models outlined above, except that this model uses the complete highway map of Mississippi, instead of only considering vertical highways. This algorithm expects three inputs outside of the parameters: the first, less accurate estimate of the affected counties, the more accurate second estimate and the final, most accurate estimate. Notice that as described, these inputs actually abstract away the hurricane's category. Instead, the category is reflected by the size of the list of the affected counties given as the respective inputs for the estimates. In order to generate these input, we used the historical hurricane data. Given the landfall and the hurricane's category, we can generate three sets of decreasing size containing neighboring counties. We can then run the simulation on these inputs. For the first 48 hours, we start to evacuate all the counties listed in the less accurate estimated, then for the second to last day we use the more accurate set of counties to pick which ones to evacuate, and finally the most accurate estimate is used for the last 24 hours.
    \par This model's main goal is to find the amount of time it takes to evacuate the affected counties, given the estimated floods. After 100 simulations with the parameters described above, we find that it consistently takes less than 50 hours to evacuate all the at-risk counties. Figures \ref{fig:stochastic_out_large} and \ref{fig:stochastic_out_small} depict the evolution of the county populations when the impending hurricane is of category 1 and 5 respectively. The hurricane in Figure~\ref{fig:stochastic_out_large} is coming from the West, and a such, the population of Jackson County initially peaks as everyone moves east and directly away from the hurricane's projected landfall, then decreases as citizens start then making their way north and out of the state. In Figure~\ref{fig:stochastic_out_small}, the hurricane comes directly from the South. There is a slight flaw in the graph in that Harrison County is represented to be lower than Jackson County, and the connectivity between the two is quite high. As such, people from Harrison County will seek refuge in Jackson County before moving North.
    \par There was also a large congestion observed in Greene County. It would seem that on their way North, citizens have trouble exiting this county. The model described in \nameref{sec:maxflow} concluded that adding highways in this area of Mississippi would drastically improve the flow of citizens. This heuristic model reflects that this is a high congestion area and supports the idea that adding roads there would increase citizen flow.
        \includegraphics[width=\linewidth]{figures/pred_hancock-crop.pdf}
        \caption{Predicted Evacuation for Landfall in Hancock County}
        \label{fig:stochastic_out_large}
        \includegraphics[width=\linewidth]{figures/pred_jackson-crop.pdf}
        \caption{Predicted Evacuation for Landfall in Jackson County}
        \label{fig:stochastic_out_small}
      \label{fig:stochastic_out}
    \par There are three main weaknesses in this model, two of which are referred to above. The first is that this model assumes that the State of Mississippi does not make accommodations for its citizens in the northern counties, meaning that once the citizens evacuate their home counties, they are left to fend for themselves. But, it does mean that evacuation is very fast, since no county officials need to be involved in organizing the evacuation. There comes a point at which the State must leverage speed of evacuation against citizen accommodations. We can only help in stating that a strategy such as the one outlined in this model will likely yield quicker evacuation times, especially in times of emergency.
    \par The other weakness is that by representing Mississippi like we have done, we prevent citizens from using smaller backroads to leave the county. This hinders their heuristic to move NorthWest in our simulations. Had we used a more accurate dataset or a more fine grid than just county geographical centers, the simulations would yield faster evacuation times, with everyone immediately going in the right direction.
